Wireless smart monitoring energy system: a proof of concept

### 2021-03-17 paper/main.tex
challenges

\section{Methods}
Throughout the development, the communication between the server and the plug has been simulated and tested using PacketSender. This tool allows the user to define the content of a packet in ASCII format, this is turned into bytes and sent to the specified address and port. For our tests, we specified the content of the packets by writing the JSON object in ASCII format, the type of packet (UDP), and the server's address and port. In particular, we tested the system's traffic and the correct handling of acknowledgments. To do this we used PacketSender to simulate the traffic generated by the plug but not the acknowledgments and Wireshark, a packet analyzer, to check the correct retransmission of packets by the server by first filtering the traffic by port and type. Finally, the entire system has been tested to check the correct transmission of acknowledgments by the plugs.

@@

\section{Challenges}
During the development, we found some issues outside the scope of this POC. Since the focus of the project wasn't on security we decided for ease of debugging to do not encrypt the packets exchanged between plugs and server. Also, we did not develop an authentication system to use the web interface.

\section{Conclusions}
In the current day and age, we have the opportunity to make smart the most mundane objects. Smart plugs can be used to monitor energy consumption, give control to the user, and help the user reach a maximum consumption goal. These plugs can reliably operate using the UDP protocol provided there are some reliability checks at the application level and can do it in a busy home WLAN where traffic can be both TCP and UDP.

>>>>

\section{Methods}
Throughout the development, the communication between the server and the plug has been simulated and tested using PacketSender. This tool allows the user to define the content of a packet in ASCII format, this is turned into bytes and sent to the specified address and port. For our tests, we specified the content of the packets by writing the JSON object in ASCII format, the type of packet (UDP), and the server address and port. In particular, we tested the system traffic and the correct handling of acknowledgments. To do this we used PacketSender to simulate the traffic generated by the plug but not the acknowledgments and Wireshark, a packet analyzer, to check the correct re-transmission of packets by the server by first filtering the traffic by port and type. Finally, the entire system has been tested to check the correct transmission of acknowledgments by the plugs.

@@

\section{Challenges}
During the development, we found some issues outside the scope of this POC. Since the focus of the project wasn't on security we decided for ease of debugging to do not encrypt the packets exchanged between plugs and server. The plugs configuration is hard-coded, for example the Wi-Fi SSID and password are stored as a char array in our C++ code. Since the EPS 8266 could switch to access point mode which means generate a new wireless network, a configuration procedure could be implemented. A device - such as a smartphone - could connect to the plug wireless network and send the SSID and password required by our home wireless settings. Those information should be stored in a persistent drive - as a micro SD card - and also, in order to not increase the POC complexity and keep the costs low, we did not implemented those procedures.
Another challenge to face is the security issue, the information send from a plug to the server are completely clear, so there is no encryption implemented. A man-in-the-middle attack could intercept those packages and change their contents. An attackers could also pose as smart plug, sending fake power readings to the server interfering with the normal operation.
\section{Conclusions}
In the current day and age, we have the opportunity to make smart the most mundane objects. Smart plugs can be used to monitor energy consumption, give control to the user, and help the user reach a maximum consumption goal. These plugs can reliably operate using the UDP protocol provided there are some reliability checks at the application level and can do it in a busy home WLAN where traffic can be both TCP and UDP.
